A Tale of Two Dictionary Learning Problems

### 2020-08-13 manuscripts/thesis/thesis.tex
new intro
l1
redid abstract

%\begin{dedication}
%\null\vfil
%\begin{center}
%I want to dedicate this to my dad. At the same time, I don't want to, because the results on display here can't be matched to the time with him I traded to produce them, and it feels as though dedicating this work to him would be measuring his life up against them. These theorems, while fundamental, are just something people might cite to feel better about the sparse coding analyses they were already going to do, anyway. And yet, despite total freedom offered by the Helen Wills Neuroscience Institute, I worried the brain was so dauntingly complex that I should feel lucky to have at last been offered a well-defined relevant problem to work on, particularly one I felt even remotely equipped to solve myself, so as to feel deserving of a Ph.D. from UC Berkeley. The fact is that whatever the eventual effect of this work on the scientific ecosystem, these truths could have waited, arrived at their own pace, via myself or another, and I could have been more present with my dad as he waited for death. I couldn't really fathom what it actually meant for him to die, to never see or talk to or hug him again, while the idea of putting original work out there, of crafting my own piece of the overwhelming puzzle that is neuroscience, was tantalizing. Let's say, then, that this dedication is the first bit of a life's work, real work void of any prestigious position or degree and the associated self-doubt and pressure, put toward passionate id	eas, solving problems I care about and that actually matter, loving and caring for others or, as my dad would say, ``being a mensch". May this dedication serve to hold myself publicly accountable to that. Also I love you, mom. I dedicate and don't dedicate this and everything to come to you, too.
%\vspace{12pt}
%\end{center}
%\vfil\null
%\end{dedication}

@@

%\begin{acknowledgements}
%I would like to thank: Chris Hillar for laying down the groundwork and vetting all of my proofs many times over. He lent me a lot of his time and expertise for little in return. Fritz Sommer for advising me and pointing me toward a tractable problem to solve. The Berkeley Student Cooperative for the amazing experience living co-operatively in Berkeley. CZ or die.
%\end{acknowledgements}

@@

\include{introduction}
\include{results}
\include{proofs}
\include{discussion}
\include{appendix}

>>>>

\begin{dedication}
\null\vfil
%\begin{center} 
I want to dedicate this to my dad. At the same time, I don't want to, because the results of this work can't be matched to the time with him I traded to produce them, and it feels as though dedicating this work to him would in some sense be measuring his life up against them. 
%These theorems, while fundamental, are just something people might cite to feel better about the sparse coding analyses they were already going to do, anyway. And yet, despite total freedom offered by the Helen Wills Neuroscience Institute, I worried the brain was so dauntingly complex that I should feel lucky to have at last been offered a well-defined relevant problem to work on, particularly one I felt even remotely equipped to solve myself, so as to feel deserving of a Ph.D. from UC Berkeley. 
The fact is that whatever may be the eventual butterfly effect of this work on the scientific ecosystem, these truths could have waited, arrived at their own pace, via myself or another, and I could have been more present with my dad as he waited patiently for a terminal cancer to take his life.

While I couldn't really fathom what it actually meant for him to die -- to never see or talk to or hug him again -- the idea of putting original work out there, of crafting my own piece of the overwhelming puzzle that is neuroscience, was alluring. Yet, I feel as though it was a feeling of inadequacy, of having to prove myself worthy of the position I had come to find myself in after a first failed project that really drove these efforts.

Let's just say, then, that this dedication is the first bit of a life's work, real work devoid of any prestigious position and the associated self-doubt, pressure, or delusions of grandeur; a life dedicated to passionate ideas, to solving problems I care about and that really matter, to loving and caring for others or, as my dad would say, to ``being a mensch". May this dedication serve to hold me publicly accountable to that. Also I love you, mom. I dedicate and don't dedicate this and everything to come to you, too.

\vspace{12pt}
%\end{center}
\vfil\null
\end{dedication}

@@

\begin{acknowledgements}
First and foremost, I would like to thank Chris Hillar for clearing the path in his original work on the dictionary problem, and for vetting all of my proofs many, many times over. He asked for nothing in return for his time, attention, and mathematical expertise, and continues to extend this generosity to anyone nerdy enough to handle it. Fritz Sommer for recommending my acceptance to Berkeley and for advising me as his student; although I eventually had to leave his lab to work on other problems of interest to me, he kindly lent me his time over the course of four years. The other members of my thesis committee: Bruno Olshausen, Mike DeWeese, and Bin Yu for their time and consideration as well. My mom would have killed me if I didn't claim this degree, so they saved my life. The graduate program managers of the Helen Wills Neuroscience Institute: Tony Leonard and Candace Groskreutz for their kindness and patience, and to the HWNI itself. The Redwood Center for Theoretical Neuroscience and the friends I made in the lab; I hope we remain in each other's lives. Finally, a heartfelt thank you to Eric Weiss for pointing me to the student co-ops, and The Berkeley Student Cooperative for the work they do providing the amazing experience of living co-operatively in Berkeley. The co-ops changed my life, and I am forever grateful -- thank you so, so much to the hundreds of talented, creative, passionate, intelligent, kind, thoughtful, absurd, inspiring people I have lived with throughout the years. The co-operative journey is only just beginning! CZ or die. 
\end{acknowledgements}

@@

\include{introduction}
\include{results}
\include{proofs}
\include{discussion}
%\include{appendix}
